### 2016-10-14 mcmc.tex
tweaks

To be explicit, imagine that you have parameters $A$ and $\phi$ and
priors $p(A)$ and $p(\phi)$.
Now imagine that instead (and very sensibly) you want to sample in
parameters $x\equiv A\,\cos(\phi)$ and $y\equiv A\,\sin(\phi)$.
What is the proper prior to apply to $x, y$?
It is not just $p(A)\,p(\phi)$ with $A=\sqrt{x^2 + y^2}$ and $\phi=\arctan(y,x)$!
That's close, though! There is a correction factor which is the
absolute value of the determinant of the Jacobian matrix
$||\frac{\dd(A,\phi}{\dd(x,y)}||$.
That is, it is the determinant of the full derivative of one set of
parameters with respect to the other.
Keep your units straight!\note{See \citet{hoggcalculus} for advice on
  how to check and test these transformations using dimensional
  analysis.}
In this particular case, the Jacobian is equal to $A$, such that
\begin{eqnarray}
  p(x,y) &=& \frac{1}{A}\,p(A)\,p(\phi)
  \quad ,
\end{eqnarray}
where $A=\sqrt{x^2 + y^2}$ and $\phi=\arctan(y,x)$.\note{Exercise to
  the reader: Do the same but with $x\equiv \sqrt{A}\,\cos(\phi)$ and
  $y\equiv\sqrt{A}\,\sin(\phi)$.}

@@

For another example, sometimes some parameters are continuous, and some
are integer (or discrete).
In this case, it is sufficient, in the context of M-H \MCMC, to make
custom proposal distributions for the integer parameters that make sure
only integers are proposed (provided care is taken to make sure that detailed
balance is preserved).
An example is to have a proposal that has some probability of
incrementing the parameter up, and an equal (ie, the same) probability
of incrementing it down.
Some samplers implicitly or explicitly assume that all parameters are
continuous\note{Once again, \citet{emcee}.}; these should be avoided
when some parameters are integer.
If you \emph{really} want to use such samplers when some parameters
are discrete, you have to put a \code{round()} or \code{int()}
operation into the likelihood and prior functions.
This makes for a stepwise-constant posterior pdf. It is a hack, but it
works in most circumstances.

>>>>

To be explicit, imagine that you have parameters $A$ and $\phi$ and
priors $p(A)$ and $p(\phi)$.
Now imagine that instead (and very sensibly) you want to sample in
parameters $x\equiv A\,\cos(\phi)$ and $y\equiv A\,\sin(\phi)$.
What is the proper prior to apply to $x, y$?
It is not just $p(A)\,p(\phi)$ with $A=\sqrt{x^2 + y^2}$ and $\phi=\arctan(y,x)$!
That's close, though! There is a correction factor which is the
absolute value of the determinant of the Jacobian matrix
$||{\dd(A,\phi)}/{\dd(x,y)}||$.
That is, it is the determinant of the full derivative of one set of
parameters with respect to the other.
Keep your units straight!\note{See \citet{hoggcalculus} for advice on
  how to check and test these transformations using dimensional
  analysis.}
In this particular case, the Jacobian is equal to $1/A$, such that
\begin{eqnarray}
  p(x,y) &=& \frac{1}{A}\,p(A)\,p(\phi)
  \quad ,
\end{eqnarray}
where $A=\sqrt{x^2 + y^2}$ and $\phi=\arctan(y,x)$.\note{Exercise to
  the reader: Do the same but with $x\equiv \sqrt{A}\,\cos(\phi)$ and
  $y\equiv\sqrt{A}\,\sin(\phi)$.}

@@

For another example, sometimes some parameters are continuous, and some
are integer (or discrete).
In this case, it is sufficient, in the context of M-H \MCMC, to make
custom proposal distributions for the integer parameters that make sure
only integers are proposed (provided care is taken to make sure that detailed
balance is preserved).
An example is to have a proposal that has some probability of
incrementing the parameter up, and an equal (that is, the same) probability
of incrementing it down.
Some samplers implicitly or explicitly assume that all parameters are
continuous\note{Once again, \citet{emcee}.}; these should be avoided
when some parameters are integer.
If you \emph{really} want to use such samplers when some parameters
are discrete, you have to put a \code{round()} or \code{int()}
operation into the likelihood and prior functions.
This makes for a stepwise-constant posterior pdf. It is a hack, but it
works in most circumstances.
